A Stopped Negative Binomial Distribution

### 2016-06-03 stat-and-prob-letters/snb-paper.tex
More paper revisions.

Consider a prototypical Phase II single-arm clinical trial in which 12 patients
are enrolled and treated. The binomial probability of success is $p=0.045$ for
each patient under the null hypothesis that treatment is not effective.
If two or more patients respond to the treatment then we reject this null
hypothesis and the trial is deemed successful at significance level of $0.1$. 
If fewer than two respond then the null is not rejected.

If all 12 patients are enrolled at once, as in the classic
design, then the sample size is 12. However, in most clinical trials the
patients are enrolled sequentially, often with one patient's outcome realized
before the next one enters the trial. In the present example, observing two
successful patients allows us to reach one endpoint so the sample required
could be as small as two. Similarly 11
observed treatment failures also ends the trial. This sampling mechanism, in
which the experiment ends as soon as a predefined endpoints is reached, is
call {\em curtailed sampling}. Under curtailed sampling the range of the
sample size for this trial is between two and 12.

@@

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients that
have been enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the fourth outcome and on in the eleventh.

\begin{figure}[t!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of a trial.
}
\label{fig:kane_viz}
\end{figure}

\begin{table}[t!]
\caption{Characteristics of the Stopped Negative Binomial Distribution}
\label{tab:snb}
\begin{center}
\begin{tabular}{|c|l|} \hline
Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
  & where $B$ is the beta function \\ \hline
MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{prop}
The distribution of the stopping time
$\argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]$
is given at (\ref{eqn:pmf}).
\end{prop}
\begin{proof}
%The proof will proceed in two parts. First, a combinatorial justification 
%will be given for the probability mass value on each element of the support. 
%Second, it will be shown that the sum of the masses over the support sums to 
%one.

@@

To show that Equation \ref{eqn:pmf} sums to one, define
\begin{align} \label{eqn:sum_proof}
R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{align}
where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
Then $R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
\citep{Olver2010} and recall this function satisfies 
$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}.
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
This completes the proof (\ref{eqn:pmf}) is is the distribution of the
stopping time and a valid probability mass
function.
\end{proof}

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters ($s$, $t$, $p$), as given. Black indicates mass contributed by hitting $s$, grey indicates
mass contributed by hitting $t$. \label{shapes.fig}}
\end{figure}

@@

There is a close connection between the tail probabilities of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to at most $s-1$ successes in a Binomial distribution. That is, the
probability that the number of successes is at least $s$ in the Binomial
model is the same that the trial succeeds (reaches $s$) in the SNB model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [Y \leq n\ |\ \text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{align*}
\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \mathcal{I}_{1-p}.
\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [Y \leq n\ |\ \text{trial succeeds}] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

To illustrate this result, let us return to our initial example, shown
in Fig.~\ref{fig:snb_bin_compare}, 
where $s=2$, $t=12$, and $p=0.045$.  The probability masses represented in 
black are equal in a) and b) as are the masses in grey.
The probability that $s$
successes are reached in the SNB process is the same as the binomial 
probability of at least two successes. Likewise, the probability that $t$ 
failures are reached in the SNB process is the same as the binomial
probability of zero or one successes.

\begin{figure}[t!]
\centering
\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.045, 2, 12) with mass contributed from 
$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 13) with
at least 2 successes (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in Equation \ref{eqn:first_sum} satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows $x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

@@

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The posterior PMF of the Stopped Negative Binomial distribution with a 
Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
f(k | s, t, \alpha, \beta) &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose k-t} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose k-t} p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose k-t} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+b-1} dp.
\end{align*}
The result follows by applying the definition of the Beta function to the integral terms.
\end{proof}

>>>>

Consider a prototypical early phase, single-arm clinical trial in which 
12 patients
are enrolled and treated. The binomial probability of success is $p=0.045$ for
each patient under the null hypothesis that treatment is not effective.
If two or more patients out of these 12 respond to the treatment then we 
reject this null
hypothesis and the treatment is deemed successful at significance 
level of $0.1$.  If fewer than two respond then the null is not rejected
and the treatment is judged as ineffective.

If all 12 patients are enrolled at once, as in the classic
design, then the sample size is 12. However, in most clinical trials the
patients are enrolled sequentially, often with one patient's outcome realized
before the next one enters the trial. In the present example, observing two
successful patients allows us to reach one endpoint so the sample required
could be as small as two. Similarly 11
observed treatment failures also ends the trial. This sampling mechanism, in
which the experiment ends as soon as a predefined endpoint is reached, is
call {\em curtailed sampling}. Under curtailed sampling the range of the
sample size for this trial is between two and 12.

@@

A hypothetical sample path is illustrated in Fig.~\ref{fig:kane_viz}.
The vertical axis denotes the number of
successful outcomes. The horizontal axis counts the number of patients 
enrolled. The horizontal and vertical boundaries represent
endpoints for the trial. In this case two successes were reached after
enrolling 10 patients: one in the fourth outcome and on in the eleventh.

\begin{figure}[t!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the trial.
}
\label{fig:kane_viz}
\end{figure}

\begin{table}[t!]
\caption{Characteristics of the Stopped Negative Binomial Distribution}
\label{tab:snb}
\begin{center}
\begin{tabular}{|c|l|} \hline
Parameters & $p$ the success probability ($0\leq p \leq 1$; $q = 1-p$) \\
           & $s$ the number of successes before stopping ($s=1, 2, ...$)\\
           & $t$ the number of failures before stopping ($t=1, 2, ...$)\\ \hline
Support & min$(s,t) \leq k \leq s+t-1$  \\ \hline
$\mathbb{P}[Y=k]$ & ${k-1 \choose s-1} p^s (1-p)^{k-s} + {k-1 \choose t-1} (1-p)^s p^{k-t}$\\ \hline
$\mathbb{P}[Y \leq k]$ & $2 - \mathcal{I}_{q}(k+1, s) - \mathcal{I}_{p}(k+1, t)$\\ 
    & where $\mathcal{I}$ is the regularized incomplete beta function.\\ \hline
Mean & $\frac{s}{p} \mathcal{I}_p(s,t) + \frac{p^{s-1} q^{t-1}}{B(s,t)} +
  \frac{t}{q} \mathcal{I}_q(t,s) + \frac{p^{t-1} q^{s-1}}{B(s,t)}$\\ 
  & where $B$ is the beta function \\ \hline
MGF & $\left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s) $\\ \hline
Posterior & 
${k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)}+
{k-1 \choose t-1} \frac{B\left(\alpha + k-t, t+\beta\right)}{B(\alpha,\beta)}$\\
& for Beta($\alpha$, $\beta$) prior. \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{prop}
The distribution of the stopping time
\begin{equation*}
Y = \argminl_k \left[X_k \geq s \cup X_k \leq k-t \right]
\end{equation*}
is given at (\ref{eqn:pmf}).
\end{prop}
\begin{proof}
%The proof will proceed in two parts. First, a combinatorial justification 
%will be given for the probability mass value on each element of the support. 
%Second, it will be shown that the sum of the masses over the support sums to 
%one.

@@

To show that (\ref{eqn:pmf}) sums to one, define
\begin{align} \label{eqn:sum_proof}
R &= \sum_{k=s}^{s+t-1} S(k, p, s) + \sum_{k=t}^{s+t-1} S(k, 1-p, t) \\
  &= \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s} + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} (1-p)^t
\end{align}
where we substitute $i=k-s$ in the first summation and $j=k-t$ in the second.
Then $R$ can be written as the cumulative distribution function of two
negative binomial distributions:
\begin{equation} \label{eqn:transformed_sum}
R = \sum_{i=0}^{t-1} {i+s-1 \choose i} p^s (1-p)^i \; + \;
\sum_{j=0}^{s-1} {j+t-1 \choose j} p^j (1-p)^t.
\end{equation}

Let $\mathcal{I}_p(s, t)$ be the {\em regularized incomplete beta function} 
\citep{Olver2010}. This function satisfies 
$\mathcal{I}_p(s, t) = 1-\mathcal{I}_{1-p}(t, s)$ \citep{Uppuluri1970}. Then
\begin{align*}
R = \sum_{i=0}^{t-1} &{i+s-1 \choose i} p^s (1-p)^i +
\sum_{j=0}^{s-1}  {j+t-1 \choose j} p^j  (1-p)^t \\
   &= 1-\mathcal{I}_p(s, t) + 1 - \mathcal{I}_{1-p}(t, s) \\
   &= 1. 
\end{align*}
This completes the proof (\ref{eqn:pmf}) is the distribution of the
stopping time and is a valid probability mass function.
\end{proof}

@@

\begin{figure}[p!]
\begin{center}
\includegraphics[width=\textwidth]{shapes.pdf}
\end{center}
\caption{Different shapes of the SNB distribution with parameters 
($p$, $s$, $t$), as given. Black indicates mass contributed by reaching
$s$ successes before $t$ failures, grey indicates
mass contributed by reaching $t$ failures first. \label{shapes.fig}}
\end{figure}

@@

There is a close connection between the tail probabilities of the SNB and the 
binomial distributions.
The probability of reaching the success endpoint in an 
SNB($p$, $s$, $t$) random variable is 
equal to the probability of at least $s$ successes in a Binomial distributed 
random variable with size $s+t-1$ and success probability $p$. 
Likewise, the probability of reaching the failure endpoint is equal
to the probability of at most $s-1$ successes in a Binomial distribution. 
That is, the probability that the number of successes is at least $s$ in the 
Binomial model is the same that the trial succeeds (reaches $s$) in the SNB 
model.
\begin{prop} \label{binomial_tail}
Let $Y$ be distributed as SNB($p$, $s$, $t$) and let 
$B$ be distributed Binomial with size $n=s+t-1$ and success probability
$p$. Then
\begin{equation}
\mathbb{P}[B \geq s] = \mathbb{P} [\text{trial succeeds}].
\end{equation}
\end{prop}
\begin{proof}
The Binomial tail probability is
\begin{align*}
\mathbb{P}[B \geq s] &= \sum_{k=s}^{s+t-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \sum_{k=0}^{s-1} {n \choose k} p^k (1-p)^{n-k} \\
  &= 1 - \mathcal{I}_{1-p}.
\end{align*}
The corresponding SNB probability is
\begin{equation*}
\mathbb{P} [\text{trial succeeds}] 
  = \sum_{k=s}^{s+t-1} {k-1 \choose s-1} p^s (1-p)^{k-s}.
\end{equation*}
Let $i=k-s$. Using the fact that
\begin{equation*}
{i+s-1 \choose s-1} = {i+s-1 \choose i}
\end{equation*}
the last summation can be rewritten as
\begin{align}
\mathbb{P} [\text{trial succeeds}] &= \sum_{i=0}^{t-1} 
  {i+s-1 \choose i} p^s (1-p)^i\\
  &= 1 - \mathcal{I}_{1-p}(t, s)
\end{align}
completing the proof.
\end{proof}

To illustrate this result, let us return to our initial example
where $s=2$, $t=12$, and $p=0.045$.  The probability masses in
Fig.~\ref{fig:snb_bin_compare} represented in 
black are equal in panels a) and b) as are the masses in grey.
The probability that $s$
successes are reached in the SNB process is the same as the binomial 
probability of at least two successes. Likewise, the probability that $t$ 
failures are reached in the SNB process is the same as the binomial
probability of zero or one successes.

\begin{figure}[t!]
\centering
\subfloat[The SNB distribution.]{\includegraphics[width=0.5\textwidth]{snb_density.pdf}}
\hfill
\subfloat[The Binomial distribution.]{\includegraphics[width=0.5\textwidth]{bin_density.pdf}}
\caption{
SNB(0.045, 2, 12) with mass contributed from 
$s$ successes (black) or $t$ failures (grey) along with Bin(0.045, 12) with
at least 2 successes (black) or fewer (grey).
}
\label{fig:snb_bin_compare}
\end{figure}

@@

The moment generating function for the SNB is calculated in a manner similar to 
that of two negative binomial distributions. 
\begin{prop} Let $Y$ be distributed SNB with parameters $p$, $s$, and $t$.
Then the moment generating function (MGF) of $Y$ is
\begin{equation} \label{eqn:mgf}
\mathbb{E}~e^{xY} = \left(\frac{p e^x}{1 - qe^x}\right)^s 
  \mathcal{I}_{1-qe^x} (s, t) + \left(\frac{qe^x}{1-pe^x}\right)^t 
  \mathcal{I}_{1-pe^x}(t, s)
\end{equation}
for $q = 1-p$ when $x \leq \min \left\{\log(1/p), \log(1/q) \right\}$.
\end{prop}
\begin{proof}
The MGF of the SNB is:
\begin{equation*}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1} {k-1 \choose k-s} p^s q^{k-s} e^{kx} 
  + \sum_{k=t}^{s+t-1} {k-1 \choose k-t} p^{k-t} q^t e^{kx}
\end{equation*}
and can be rewritten as:
\begin{equation} \label{eqn:first_sum}
\mathbb{E}~e^{xY} = \sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} 
  + \sum_{k=t}^{s+t-1}{k-1 \choose k-t} (qe^x)^t (pe^x)^{k-t}.
\end{equation}
The first summation in (\ref{eqn:first_sum}) satisfies
\begin{align*}
\sum_{k=s}^{s+t-1}{k-1 \choose k-s} (pe)^{sx} (qe^x)^{k-s} &= 
  \left(\frac{pe^x}{1 - qe^x}\right)^s \ \ \sum_{k=s}^{s+t-1} {k-1 \choose k-s} 
    (qe^x)^{k-s} (1-qe^x)^s \\
  &= \left(\frac{pe^x}{1 - qe^x}\right)^s \mathcal{I}_{qe^x}(s, t).
\end{align*}
Since the incomplete beta function's subscript parameter has support on zero 
to one, we have $qe^x \leq 1$. This also shows we must restrict
$x \leq -\log(q)$.
A similar expression can be derived for the second summation in 
Equation \ref{eqn:first_sum} and results in
the constraint $x \leq -\log(p)$.
\end{proof}

@@

Let us consider a Bayesian analysis with a Beta prior distribution on $p$.
\begin{prop} \label{prop:bayesian}
The posterior probability mass function of the Stopped Negative Binomial 
distribution with a 
Beta($\alpha$, $\beta$) prior is
\begin{align} \label{eqn:posterior}
\mathbb{P} [Y = k | s, t, \alpha, \beta ] &= 
  {k-1 \choose s-1} \frac{B\left(\alpha+s, k-s+\beta \right)}{B(\alpha, \beta)} 
    \ I_{\{s \leq k \leq s+k-1\}} \ + \nonumber \\
  & {k-1 \choose t-1} 
    \frac{B\left(\alpha + k - t, t+\beta\right)}{B(\alpha, \beta)} 
    \ I_{\{t \leq k \leq s+k-1\}}.
\end{align}
\end{prop}
\begin{proof}
For notational simplicity, assume that $k$ is in the range
$\min(s,t) \leq k \leq s+t-1$. When 
this is not the case appropriate terms should be removed as dictated by the indicator functions.
\begin{align*}
f(k | s, t, \alpha, \beta) = \frac{1}{B(\alpha, \beta)} & \int_0^1 {k-1 \choose s-1} p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} + \\
 & {k-1 \choose t-1} p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp \\
= \frac{1}{B(\alpha, \beta)}  {k-1 \choose s-1} & \int_0^1  p^{\alpha +s -1} \left(1-p\right)^{k-s+\beta-1} dp + \\
 & \frac{1}{B(\alpha, \beta)} {k-1 \choose t-1} \int_0^1  p^{k-t+\alpha-1}\left(1-p\right)^{t+\beta-1} dp.
\end{align*}
The result in (\ref{eqn:posterior}) follows by the definition of the 
Beta function.
\end{proof}

To better understand the posterior distribution, let us once again return to 
our example where $s=2$, $t=12$. However, we assume $p$ is a Beta distribution
with shape parameters $\alpha = 0.045 c$ and 
$\beta = 0.955 c$ where $c$ will take values between one and 10. 
This parameterization allows us to examine the relationship between 
uncertainty in $p$ and the SNB.

\begin{figure}[t!]
\centering
\subfloat[The expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[The variance of the sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The expected size and variance of the posterior SNB varying the uncertainty 
in the prior distribution of $p$. A larger value of $c$
corresponds to a Beta prior with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}

Fig.~\ref{fig:bayesian-sample-size} shows how the expected sample size
and variance of the posterior SNB. As the certainty increases the 
posterior values converge to corresponding expected values with 
$p = 0.045$. It may be noted that convergence is relatively slow. This
is because when $c$ is small, the prior is heavily skewed. As $c$ increases
this effect is minimized.

Slow convergence in the mean will occur when $p$ takes values close to 
zero or one. In these cases it may be preferable to select shape paramters
based on the median or mode of the beta distribution. These values
will vary less in the uncertainty of the distribution.
